---
title: "Final Project"
author: "Cynthia Petrossian"
date: "5/9/2022"
output: 
  prettydoc::html_pretty:
    toc: TRUE
    theme: cayman
    
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      fig.path = "../Outputs/")
```

# Load Libraries
```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggpubr)
library(gganimate)
library(gifski)
library(prettydoc)
library(magick)
```

# Introduction
The integrity of organs and bodily tissues relies on adult stem cells (ASCs). ASCs divide to replace any cells lost to damage, disease, or regular turnover. There are many factors that control the differentiation and division of ASCs, notably master regulators (MRs) and their downstream targets.  

```{r}
image_read(here("Final_Project", "Images", "iscs.jpg"))  #Read in ISC image
```

While the effects of manipulating MRs on ASCs have been well characterized, and many of their potential downstream targets identified, validation of these targets and a mechanistic explanation of how they function within ASCs is still largely lacking.

To explore this, we use the model system *Drosophila melanogaster*, the fruit fly. We specifically study intestinal stem cells (ISCs) in the posterior midgut region of the fly. 
```{r}
image_read(here("Final_Project", "Images", "modelsystem.png"))  #Read in ISC image
```

We will focus on two well-known MRs of ISCs in *Drosophila*, Escargot (Esg) and STAT. While previous work in our lab has validated several targets regulated by Esg, STAT, or both, the role of these targets themselves in ISC regulation has not been investigated.

We will attempt to elucidate the roles of these targets by manipulating their genetic expression and observing the effects of this manipulation on ISC morphology. 

For this project, we will focus on the downstream genetic target Indy. We genetically manipulated the expression of Indy by overexpressing it in *Drosophila* ISCs and visualized the intestinal cells through immunofluorescence microscopy. 

We used Cell Profiler, a "free, open-source software designed to ... quantitatively measure phenotypes from thousands of images automatically" to generate numerical data of the phenotypes observed. This data was used for the following plots. 

# Indy Overexpression Phenotypes

```{r}
image_read(here("Final_Project", "Images", "indyoe.png"))  #Read in ISC image
```

We visually saw an increase in the number of red enteroendocrine (EE) cells relative to the number of green progenitor cells in Indy OE.  

I will read in the data of EE and progenitor cell counts and plot it.

# Increase in Enteroendrocrine Counts
```{r eeprog}
cells <- read_csv(here("Final_Project", "Data", "EE_Progenitor_Counts", "ee_progenitor_counts.csv")) #read in data
view(cells) #view data

cells %>% 
  ggplot(aes(x = group, #create ggplot, set x axis as experimental group
             y = ratio, #set y axis as ratio
             color = group))+ #set color as group
  geom_boxplot()+ #geom boxplot
  geom_point(position = position_jitterdodge())+ #add jittered points
  labs(title = "EE Counts Relative to Progenitor Counts in Indy OE (20117)", #add title
       y = "# of EE Cells / # of Progenitor Cells", #add y axis label
       x = "Experimental Group")+ #add x axis label
  guides(color = FALSE)+#remove legend
   geom_signif(comparisons = list(c("Ctl", "Indy OE")), #determine significance
              map_signif_level=TRUE, #show significance on plot
              color = "black")  #change bracket color to black
```

# Indy Overexpression Phenotypes
```{r}
image_read(here("Final_Project", "Images", "indyoe.png"))  #Read in ISC image
```
We also noticed that it appeared as if blue enterocytes (ECs) were more spread out from each other in the Indy OE compared to control. We used CellProfiler to quantify the distances between ECs in control and Indy OE guts. We automatically detected EC cells from microscopy images using CellProfiler, and then extracted measurements of the distances between each EC and its first and second closest neighboring EC.

First I will read in the data and wrangle it into an appropriate structure for plotting.  

# 20117 Data Wrangling
```{r}
indy1ctl <- read_csv(here("Final_Project", "Data", "20117",  "DAPIcells_Filtered_DAPI_nuclei_ctl.csv")) #read in data
indy1exp <- read_csv(here("Final_Project", "Data", "20117",  "DAPIcells_Filtered_DAPI_nuclei_exp.csv")) #read in data

indy1ctl <- indy1ctl %>% 
  clean_names() #clean column names with janitor package
indy1ctl #view indy1ctl

indy1exp <- indy1exp %>% 
  clean_names() #clean column names with janitor package
indy1exp #view indy1exp

indy1ctl <- mutate(.data = indy1ctl,
                    metadatagroup = "Ctl",
                    line = "20117") #add new column for metadatagroup to indy1ctl dataframe - have it say "Ctl"

indy1exp <- mutate(.data = indy1exp,
                    metadatagroup = "Indy OE",
                    line = "20117") #add new column for metadatagroup to indy1exp  - have it say "Indy OE"


combinedindy1 <- bind_rows(indy1ctl, indy1exp) #stack the two dataframes on top of each other
combinedindy1 #view combined dataframes

means_combinedindy1 <- combinedindy1 %>% #create new df
  group_by(metadatagroup, metadata_gut, line) %>% #group by experimental group and gut number
  summarise(first_distance_mean = mean(neighbors_first_closest_distance_expanded), #get means for first distance
            second_distance_mean = mean(neighbors_second_closest_distance_expanded)) %>% #get means for second distance
  mutate(microns_first = first_distance_mean/6.20, #add new column called "microns_first" - convert distances from pixels to microns by dividing by 6.2
         microns_second = second_distance_mean/6.20) #add new column called "microns_second" - convert distances from pixels to microns by dividing by 6.2
view(means_combinedindy1) #viewnew dataframe

rfindy1 <- means_combinedindy1 %>% #create new dataframe
  rowwise() #use rowwise to allow for computing of dataframe on row by row basis
rfindy1 #view dataframe
```

# 21412 Data Wrangling
```{r}
indy2ctl <- read_csv(here("Final_Project", "Data", "21412", "DAPIcells_Filtered_DAPI_nuclei_ctl.csv")) #read in data
indy2exp <- read_csv(here("Final_Project", "Data", "21412",  "DAPIcells_Filtered_DAPI_nuclei_exp.csv")) #read in data

indy2ctl <- indy2ctl %>% 
  clean_names() #clean names using janitor package
indy2ctl #view df

indy2exp <- indy2exp %>% 
  clean_names() #clean names using janitor package
indy2exp #view df

indy2ctl <- mutate(.data = indy2ctl, 
                    metadatagroup = "Ctl",
                    line = "21412") #add new column to df called "metadatagrop" and set as "Ctl"

indy2ctl
indy2exp <- mutate(.data = indy2exp,
                    metadatagroup = "Indy OE",
                    line = "21412") #add new column to df called "metadatagrop" and set as "Indy OE"

combinedindy2 <- bind_rows(indy2ctl, indy2exp) #stack dfs on top of each other
view(combinedindy2) #view new df

means_combinedindy2 <- combinedindy2 %>% #create new df
  group_by(metadatagroup, metadata_gut, line) %>% #group by metaadatagroup and gut number
  summarise(first_distance_mean = mean(neighbors_first_closest_distance_expanded), #get first distance means
            second_distance_mean = mean(neighbors_second_closest_distance_expanded)) %>% #get second distance means
  mutate(microns_first = first_distance_mean/6.20, #add new column called "microns_first" to convert pixels to microns by dividing by 6.2
         microns_second = second_distance_mean/6.20) #add new column called "microns_second" to convert pixels to microns by dividing by 6.2
view(means_combinedindy2) #view df

rfindy2 <- means_combinedindy2 %>% #create new df
  rowwise() #use rowwise to allow for computing dataframe on row by row basis
rfindy2 #view df
```

I will them combine the two dataframes to use in one type of plot.  

# Data Wrangling of Both 20117 & 21412
```{r}
combinedindy <- bind_rows(means_combinedindy1, means_combinedindy2) #stack dfs on top of each other
view(combinedindy) #view new df

rfindy <- combinedindy %>% 
  rowwise() #use rowwise to allow for computing dataframe on row by row basis
rfindy #view df
```

I will now begin plotting.   


# First Distance Means Plot (Animated)
```{r firstdistancecombined}
first <- combinedindy %>% #create ggplot from this data
  ggplot(aes(x = metadatagroup, #set x axis as metadatagroup
             y = microns_first, #set y axis as microns
             color = metadatagroup))+ #set color as metadatagroup
  geom_boxplot()+ #geom is boxplot
  geom_point(position = position_jitterdodge())+
  labs(x = "Experimental Group", #add x axis label
       y = "Average of Means (micrometers)")+ #add y axis label
  guides(color = FALSE)+ #remove legend
transition_states(line, #the thing you are animating by
                    transition_length = 2, #relative length of transition (length between transitions)
                    state_length = 1)+ #The length between end and beginning of gif
  ease_aes("sine-in-out")+
  ggtitle("First Distance Means of Line {closest_state}")

  animate(first, height=400, width=600, renderer=gifski_renderer())
  anim_save("combinedmeans.gif", first)
```

# 20117 First Distance Means Plot
```{r 20117firstdistance}
means_combinedindy1 %>% #create ggplot from this data
  ggplot(aes(x = metadatagroup, #set x axis as metadatagroup
             y = microns_first, #set y axis as microns
             color = metadatagroup))+ #set color to metedatagroup
  geom_boxplot()+ #geom is boxplot
  geom_point(position = position_jitterdodge())+
  labs(title = "First Distance Means in 20117", #add title
       y = "First Distance Means (micrometers)", #add y axis label
       x = "Experimental Group")+ #add x axis label
  guides(color = FALSE)+ #remove legend
  geom_signif(comparisons = list(c("Ctl", "Indy OE")), #get significance
              map_signif_level=TRUE, #add significance to plot
              color = "black")  #change bracket color to black
```

# 21412 First Distance Means Plot
```{r 21412firstdistance}
means_combinedindy2 %>% #create ggplot from this data
  ggplot(aes(x = metadatagroup, #set x axis as metadatagroup
             y = microns_first, #set y axis as microns
             color = metadatagroup))+ #set color as metadatagroup
  geom_boxplot()+ #geom is boxplot
  geom_point(position = position_jitterdodge())+ #add jittered points
  labs(title = "First Distance Means in 21412", #add title
       y = "First Distance Means (micrometers)", #add y axis label
       x = "Experimental Group")+ #add x axis label
  guides(color = FALSE)+ #remove legend
  geom_signif(comparisons = list(c("Ctl", "Indy OE")), #get significance
              map_signif_level=TRUE, #add significance to plot
              color = "black")  #change bracket color to black
```

# Sum of First and Second Distance Means Plot (Animated)
```{r combinedsumplot}
rfsumindy <- rfindy %>% #create new dataframe
  mutate(sum = sum(c(microns_first, microns_second)))

sum <- rfsumindy %>% #create ggplot from this data
  ggplot(aes(x = metadatagroup, #set x axis as metadatagroup
             y = microns_first, #set y axis as microns
             color = metadatagroup))+ #set color as metadatagroup
  geom_boxplot()+ #geom is boxplot
  geom_point(position = position_jitterdodge())+
  labs(x = "Experimental Group", #add x axis label
       y = "Sum of Means (micrometers)")+ #add y axis label
  guides(color = FALSE)+ #remove legend
transition_states(line, #the thing you are animating by
                    transition_length = 2, #relative length of transition (length between transitions)
                    state_length = 1)+ #The length between end and beginning of gif
  ease_aes("sine-in-out")+
  ggtitle("Sum of First and Second Means of Line {closest_state}")

  animate(sum, height=400, width=600, renderer=gifski_renderer())
  anim_save("combinedsum.gif", sum)
```

# 20117 Sum of First and Second Distance Means Plot
```{r 20117sumplot}
rfsumindy1 <- rfindy1 %>% #create new dataframe
  mutate(sum = sum(c(microns_first, microns_second))) #add new colmn to rowwise dataframe - calclate sum of first and second distances per gut
rfsumindy1

rfsumindy1 %>% #create ggplot from this data
  ggplot(aes(x = metadatagroup, #set x axis as metadatagroup
             y = sum, #set y axis as sum
             color = metadatagroup))+ #set color as metadatagroup
  geom_boxplot()+ #geom is boxplot
  geom_point(position = position_jitterdodge())+ #add jittered points
  labs(title = "Sum of First and Second Means 20117", #add title
       x = "Experimental Group", #add x axis label
       y = "Sum of Means (micrometers)")+ #add y axis label
  guides(color = FALSE)+ #remove legend
  geom_signif(comparisons = list(c("Ctl", "Indy OE")), #get significance
              map_signif_level=TRUE, #add significance to plot
              color = "black")  #change bracket color to black
```

# 21412 Sum of First and Second Distance Means Plot
```{r 21412sumplot}
rfsumindy2 <- rfindy2 %>% #create new df
  mutate(sum = sum(c(microns_first, microns_second))) #add column for sum of first and second distances
rfsumindy2 #view new df

rfsumindy2 %>% #create ggplot from this data
  ggplot(aes(x = metadatagroup, #set x axis to metadatagroup
             y = sum, #set y axis to sum
             color = metadatagroup))+ #set color to metadatagroup
  geom_boxplot()+ #geom is boxplot
  geom_point(position = position_jitterdodge())+ #add jittered points
  labs(title = "Sum of First and Second Means 21412", #add title
       x = "Experimental Group", #add x axis label
       y = "Sum of Means (micrometers)")+ #add y axis label
  guides(color = FALSE)+ #remove legend
  geom_signif(comparisons = list(c("Ctl", "Indy OE")), #get significance
              map_signif_level=TRUE, #add signifcance to plot
              color = "black")  #change bracket color to black
```

# Average of First and Second Distance Means Plot (Animated)
```{r combinedavgplot}
rfavgindy <- rfindy %>% #create new dataframe
  mutate(mean = mean(c(microns_first, microns_second)))
rfavgindy

average <- rfavgindy %>% #create ggplot from this data
  ggplot(aes(x = metadatagroup, #set x axis as metadatagroup
             y = microns_first, #set y axis as microns
             color = metadatagroup))+ #set color as metadatagroup
  geom_boxplot()+ #geom is boxplot
  geom_point(position = position_jitterdodge())+
  labs(x = "Experimental Group", #add x axis label
       y = "Average of Means (micrometers)")+ #add y axis label
  guides(color = FALSE)+ #remove legend
transition_states(line, #the thing you are animating by
                    transition_length = 2, #relative length of transition (length between transitions)
                    state_length = 1)+ #The length between end and beginning of gif
  ease_aes("sine-in-out")+
  ggtitle("Average of First and Second Means of Line {closest_state}")
animate(average, height=400, width=600, renderer=gifski_renderer())
anim_save("combinedavg.gif", average)
```

# 20117 Average of First and Second Distance Means Plot
```{r 20117avgplot}
rfavgindy1 <- rfindy1 %>% #create new dataframe
  mutate(avg = mean(c(microns_first, microns_second))) #calculate average of first and second distances
rfavgindy1 #view dataframe

rfavgindy1 %>% #create ggplot from this data
  ggplot(aes(x = metadatagroup, #set x axis as metadatagroup
             y = avg, #set y axis as average
             color = metadatagroup))+ #set color as metadatagroup
  geom_boxplot()+ #geom is boxplot
  geom_point(position = position_jitterdodge())+ #add jittered points
  labs(title = "Average of First and Second Means 20117", #add title
       x = "Experimental Group", #add x axis label
       y = "Average of Means (micrometers)")+ #add y axis label
  guides(color = FALSE)+ #remove legend
  geom_signif(comparisons = list(c("Ctl", "Indy OE")), #get significance
              map_signif_level=TRUE, #add significance to plot
              color = "black")  #set bracket color to black
```
# 21412 Average of First and Second Distance Means Plot
```{r 21412avgplot}
rfavgindy2 <- rfindy2 %>% #create new df
  mutate(avg = mean(c(microns_first, microns_second))) #add new column for average of first and second distance
rfavgindy2 #view new df

rfavgindy2 %>% #create ggpot from this data
  ggplot(aes(x = metadatagroup, #set x axis to metadatagroup
             y = avg, #set y axis to avg
             color = metadatagroup))+ #set color to metadatagroup
  geom_boxplot()+ #geom is boxplot
  geom_point(position = position_jitterdodge())+ #add jittered points
  labs(title = "Average of First and Second Means 21412", #add title
       x = "Experimental Group", #add x axis label
       y = "Average of Means (micrometers)")+ #add y axis label
  guides(color = FALSE)+ #remove legend
  geom_signif(comparisons = list(c("Ctl", "Indy OE")), #get significance
              map_signif_level=TRUE, #dd singificance to plot
              color = "black")  #change bracket color to black
```